Capital Budget Funding in Toronto

### 2024-09-26 paper/paper.qmd
tried making data points red

# Format numeric columns with conditional formatting for LaTeX
merged_table_2 <- merged_table_select |>
  mutate(
    # Format Per Capita Funding as dollars
    per_capita_funding = dollar(per_capita_funding),
    `Median total income of households in 2020 ($)` = dollar(as.numeric(`Median total income of households in 2020 ($)`)),
    `Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)` = paste0(round(as.numeric(`Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)`) * 100, 2), "%"),
    `Bachelor’s degree or higher per capita` = paste0(round(as.numeric(`Bachelor’s degree or higher per capita`) * 100, 2), "%"),
    `Total visible minority population per capita` = paste0(round(as.numeric(`Total visible minority population per capita`) * 100, 2), "%")
  )

# Create the table with kable and apply enhanced styling
kable(
  merged_table_2, 
  escape = TRUE,                   
  col.names = c(
    "Ward Number", 
    "Per Capita Funding", 
    "Household Median Income", 
    "Low Income Prevalence", 
    "Bachelor’s Degree or Higher", 
    "Visible Minority Population"
  )
) |>
  kable_styling(
    bootstrap_options = c("condensed"),  
    full_width = FALSE,  
    position = "left",  
    font_size = 7 
  ) |>
  column_spec(1, width = "2cm") |>
  column_spec(2:3, width = "3cm") |>
  column_spec(4, width = "2cm") |>
  column_spec(5:6, width = "3cm")

@@

- **Ward 2** has a household median income of **$100,000** and receives **$4,166.43** per capita in funding.
- **Ward 10**, with a median income of **$89,000**, receives **$4,127.10** per capita.
- Conversely, **Ward 22**, with a lower median income of **$77,000**, receives only **$192.44** per capita.

**Figure 1** illustrates this relationship, showing a general upward trend. While there are exceptions—such as **Ward 13**, which has a median income of **$65,000** but the highest per capita funding—the overall pattern indicates that wealthier wards receive more capital investment.

@@

- **Ward 22** has a low-income prevalence of **15%** and receives **$192.44** per capita.
- **Ward 5**, with **15%** low-income prevalence, receives **$867.02** per capita.
- In contrast, **Ward 2** has a low-income prevalence of **8%** and receives **$4,166.43** per capita.

@@

- **Ward 10**, where **60.97%** of residents have at least a bachelor's degree, receives **$4,127.10** per capita.
- **Ward 11**, with **56.07%** of residents holding a bachelor's degree or higher, receives **$1,637.81** per capita.
- Conversely, **Ward 7**, with only **14.67%** educational attainment at this level, receives **$852.05** per capita.

@@

- **Ward 22** has a visible minority population of **82.31%** and receives **$192.44** per capita.
- **Ward 23**, with a visible minority population of **92.18%**, receives **$1,829.15** per capita.
- In contrast, **Ward 2**, with a visible minority population of **31.75%**, receives **$4,166.43** per capita.

>>>>

# Format numeric columns with conditional formatting for LaTeX
merged_table_2 <- merged_table_select |>
  mutate(
    # Format Per Capita Funding as dollars
    per_capita_funding = dollar(per_capita_funding),
      `Median total income of households in 2020 ($)` = paste0(
      dollar(as.numeric(`Median total income of households in 2020 ($)`)),
      ifelse(as.numeric(`Median total income of households in 2020 ($)`) <= 84000, "**", "")
    ),
    `Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)` = paste0(round(as.numeric(`Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)`) * 100, 2), "%"),
    `Bachelor’s degree or higher per capita` = paste0(round(as.numeric(`Bachelor’s degree or higher per capita`) * 100, 2), "%"),
    `Total visible minority population per capita` = paste0(round(as.numeric(`Total visible minority population per capita`) * 100, 2), "%")
  )

# Create the table with kable and apply enhanced styling
kable(
  merged_table_2, 
  escape = TRUE,                   
  col.names = c(
    "Ward Number", 
    "Per Capita Funding", 
    "Household Median Income", 
    "Low Income Prevalence", 
    "Bachelor’s Degree or Higher", 
    "Visible Minority Population"
  )
) |>
  kable_styling(
    bootstrap_options = c("condensed"),  
    full_width = FALSE,  
    position = "left",  
    font_size = 7 
  ) |>
  column_spec(1, width = "2cm") |>
  column_spec(2:3, width = "3cm") |>
  column_spec(4, width = "2cm") |>
  column_spec(5:6, width = "3cm")
```

Wards with a **\*\*** indicate a **Household Median Income** of \$84,000 or less.

@@

-   **Ward 2** has a household median income of **\$100,000** and receives **\$4,166.43** per capita in funding.
-   **Ward 10**, with a median income of **\$89,000**, receives **\$4,127.10** per capita.
-   Conversely, **Ward 22**, with a lower median income of **\$77,000**, receives only **\$192.44** per capita.

**Figure 1** illustrates this relationship, showing a general upward trend. While there are exceptions—such as **Ward 13**, which has a median income of **\$65,000** but the highest per capita funding—the overall pattern indicates that wealthier wards receive more capital investment.

@@

-   **Ward 22** has a low-income prevalence of **15%** and receives **\$192.44** per capita.
-   **Ward 5**, with **15%** low-income prevalence, receives **\$867.02** per capita.
-   In contrast, **Ward 2** has a low-income prevalence of **8%** and receives **\$4,166.43** per capita.

@@

-   **Ward 10**, where **60.97%** of residents have at least a bachelor's degree, receives **\$4,127.10** per capita.
-   **Ward 11**, with **56.07%** of residents holding a bachelor's degree or higher, receives **\$1,637.81** per capita.
-   Conversely, **Ward 7**, with only **14.67%** educational attainment at this level, receives **\$852.05** per capita.

@@

-   **Ward 22** has a visible minority population of **82.31%** and receives **\$192.44** per capita.
-   **Ward 23**, with a visible minority population of **92.18%**, receives **\$1,829.15** per capita.
-   In contrast, **Ward 2**, with a visible minority population of **31.75%**, receives **\$4,166.43** per capita.
